GEE models

### 2017-10-16 doc/manuscript.Rmd
clarified that diabetes cases were not used only at/after visit
clarified med us

For the primary analysis, generalized estimating equation (GEE) models 
[@Zeger1986a] were used to determine the longitudinal associations between the 
outcome variables and the predictor variables. Given the longitudinal design an
auto-regressive of order 1 (AR1) correlation matrix was chosen for the GEE
models, though other matrices (eg. exchangeable) had similar fit (data not
shown). GEE is well suited to longitudinal cohort studies given it's capacity to
handle any missed visits and for the correlation of variables between time points. 
See the ESM Methods for more detail on GEE.
The predictor variables and continuous covariates were
scaled (mean centered and standardised). The NEFA variables were classified as 
*time-independent* (held constant) as they were measured only at the baseline 
visit, while the outcome variables were set as *time-dependent*. No imputation
was conducted on missing values and incident diabetes cases were excluded from 
the analysis. A GEE model is applied for each combination of outcome and NEFA (as
either mol% or nmol/ml).

@@

For this study, there were `r female$npct` females and `r european$npct` had
European-ancestry (see `r cite_t("basic")`). The mean follow-up time was 
`r calc_followup_time()` years, where `r calc_pct_full_visits()`% of 
participants attended all three visits.
The average age of the participants at baseline was `r aide::average(ds0yr$Age)`
years (`r aide::stddev(ds0yr$Age)` SD) and the average BMI was
`r aide::average(ds0yr$BMI)` kg/m^2^ (`r aide::stddev(ds0yr$BMI)` SD). Most of the 
participants, `r famhist$npct`, had a family history of diabetes. Insulin
sensitivity and beta cell function measures showed a significant median
decline from baseline to 6-years of between `r chg_outcomes$chg` 
(p`r chg_outcomes$p`, based on a time-adjusted GEE model) in this analysis.
Consistent with this decline, there were `r calc_incid_dysgly("ConvertDM")`
and `r calc_incid_dysgly("ConvertPreDM")`incident cases of diabetes and
pre-diabetes (IFG and IGT), respectively, over the 6-years. Incident diabetes
cases were excluded from further GEE analyses. There were `r meds_use[1]` and
`r meds_use[2]` participants who used lipid and cholesterol modifying medications
at the baseline and 6-year visit, respectively. Medication use did not influence
NEFA composition, the outcomes, or the GEE models (data not shown).

>>>>

For the primary analysis, generalized estimating equation (GEE) models 
[@Zeger1986a] were used to determine the longitudinal associations between the 
outcome variables and the predictor variables. Given the longitudinal design an
auto-regressive of order 1 (AR1) correlation matrix was chosen for the GEE
models, though other matrices (eg. exchangeable) had similar fit (data not
shown). GEE is well suited to longitudinal cohort studies given it's capacity to
handle any missed visits and for the correlation of variables between time points. 
See the ESM Methods for more detail on GEE.
The predictor variables and continuous covariates were
scaled (mean centered and standardised). The NEFA variables were classified as 
*time-independent* (held constant) as they were measured only at the baseline 
visit, while the outcome variables were set as *time-dependent*. No imputation
was conducted on missing values and incident diabetes cases were excluded from 
the analysis (follow-up data were included until date of diagnosis). A GEE model
is applied for each combination of outcome and NEFA (as either mol% or nmol/ml).

@@

For this study, there were `r female$npct` females and `r european$npct` had
European-ancestry (see `r cite_t("basic")`). The mean follow-up time was 
`r calc_followup_time()` years, where `r calc_pct_full_visits()`% of 
participants attended all three visits.
The average age of the participants at baseline was `r aide::average(ds0yr$Age)`
years (`r aide::stddev(ds0yr$Age)` SD) and the average BMI was
`r aide::average(ds0yr$BMI)` kg/m^2^ (`r aide::stddev(ds0yr$BMI)` SD). Most of the 
participants, `r famhist$npct`, had a family history of diabetes. Insulin
sensitivity and beta cell function measures showed a significant median
decline from baseline to 6-years of between `r chg_outcomes$chg` 
(p`r chg_outcomes$p`, based on a time-adjusted GEE model) in this analysis.
Consistent with this decline, there were `r calc_incid_dysgly("ConvertDM")`
and `r calc_incid_dysgly("ConvertPreDM")`incident cases of diabetes and
pre-diabetes (IFG and IGT), respectively, over the 6-years. Incident diabetes
cases were excluded from GEE analyses (follow-up data were included until date
of diagnosis). Certain medications (e.g. lipid medications such as statins or
fibrates) could influence the association between NEFA and the outcomes.
There were `r meds_use[1]` and `r meds_use[2]` participants who used lipid and
cholesterol modifying medications at the baseline and 6-year visit,
respectively. Use of these medications did not influence NEFA composition, the
outcomes, or the GEE models (data not shown).
